---
title: "GLMER Gene Analysis"
author: "Tim Neher"
date: "12/10/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, results=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(lme4)
library(emmeans)
library(performance)
library(sjPlot)
```

```{r, results=FALSE}
qpcr_r <- read.csv("qpcr_r.csv")

qpcr_r$integrator <- as.factor(qpcr_r$integrator)
qpcr_r$biorep <- as.factor(qpcr_r$biorep)
qpcr_r$techrep <- as.factor(qpcr_r$techrep)
qpcr_r$farm <- as.factor(qpcr_r$farm)
```

```{r, results=FALSE}
rna <- subset(qpcr_r, gene == "16S")
ermb <- subset(qpcr_r, gene == "ermB")
tetm <- subset(qpcr_r, gene == "tetM")
```

# Integrator and Production System Effects GLMER

## ermB
### Data curation
Keeps the manure samples and farms with antibiotics dosed in feed and water (n farms=47)
```{r}
ermb_1 <- subset(ermb, matrix == 'manure')
ermb_1 <- ermb_1 %>%
  filter(!antibiotic == ("no"))
```

### Model setup
```{r}
modelermb <- glmer(copies.g ~ type*integrator + (1|farm:biorep), data = ermb_1, family = Gamma(link = "log"))
modelermb

model_performance(modelermb)
```

### ermB EMMEANS for integrator and production system
```{r}
modelermb.emm <- emmeans(modelermb, c("integrator", "type"))
modelermb.emm
pairs(modelermb.emm)

modelermb.emm1 <- emmeans(modelermb, c("integrator"))
modelermb.emm1

modelermb.emm2 <- emmeans(modelermb, c("type"))
modelermb.emm2

joint_tests(modelermb.emm)

joint_tests(modelermb.emm, by = "type")
joint_tests(modelermb.emm, by = "integrator")

emm_ermb_i.t <- emmeans(modelermb, pairwise ~ integrator | type)
emm_ermb_i.t

emm_ermb_t.i <- emmeans(modelermb, pairwise ~ type | integrator)
emm_ermb_t.i
```


### ermB estimated marginal means plot for integrator and production system
```{r}
ermb <- emmip(modelermb, integrator ~ type, CIs = TRUE, PIs = FALSE, ylab = "Estimated Marginal Mean (Ln Copies/Gram)",
              xlab = "Production System", tlab = "Integrator", linearg = list(linetype = "dashed"), CIarg = list(lwd = 2, alpha = 1),
              dotarg = list(size=10, shape=3))

ermb <- ermb + scale_color_manual(values=c("#F97066", "#2586E8")) + font_size(labels.x = "12", axis_title.x = "12", axis_title.y = "12", labels.y = "12")

ermb <- ermb + scale_y_continuous(breaks = c(12, 13, 14, 15, 16, 17, 18))

ermb <- ermb + expand_limits(y = (c(12, 18)))

ermb
```

## tetM integrator and production system effects
### Data curation
Keeps the manure samples and farms with antibiotics dosed in feed and water (n farms=47)
```{r}

tetm_1 <- subset(tetm, matrix == 'manure')
tetm_1 <- tetm_1 %>%
  filter(!antibiotic == "no")
```

### Model setup
```{r}
modeltetm <- glmer(copies.g ~ type*integrator + (1|farm:biorep), data = tetm_1, family = Gamma(link = "log"))
modeltetm

model_performance(modeltetm)
```


### tetM EMMEANS for integrator and producion system
```{r}
modeltetm.emm <- emmeans(modeltetm, c("integrator", "type"))
modeltetm.emm
pairs(modeltetm.emm)

modeltetm.emm1 <- emmeans(modeltetm, c("integrator"))
modeltetm.emm1
modeltetm.emm2 <- emmeans(modeltetm, c("type"))
modeltetm.emm2

joint_tests(modeltetm.emm)

joint_tests(modeltetm, by = "type")
joint_tests(modeltetm, by = "integrator")

emm_tetm_i.t <- emmeans(modeltetm, pairwise ~ integrator | type)
emm_tetm_i.t

emm_tetm_t.i <- emmeans(modeltetm, pairwise ~ type | integrator)
emm_tetm_t.i
```

### tetM estimated marginal means plot for integrator and production system
```{r}
tetm <- emmip(modeltetm, integrator ~ type, CIs = TRUE, PIs = FALSE, ylab = "Estimated Marginal Mean (Ln Copies/Gram)",
              xlab = "Production System", tlab = "Integrator", linearg = list(linetype = "dashed"), CIarg = list(lwd = 2, alpha = 1),
              dotarg = list(size=10, shape=3)) 

tetm <- tetm + scale_color_manual(values=c("#F97066", "#2586E8")) + font_size(labels.x = "12", axis_title.x = "12", axis_title.y = "12", labels.y = "12")

tetm <- tetm + scale_y_continuous(breaks = c(12, 13, 14, 15, 16, 17, 18))

tetm <- tetm + expand_limits(y = (c(12, 18)))

tetm
```

# Antibiotic Effects GLMER

## ermB
### Data curation
Keeps the ermB gene and removes the integrator 1 and Gilt from the dataset.
```{r}
ermb <- subset(qpcr_r, gene == "ermB")
ermb_ant <- ermb %>%
  filter(integrator == ("2") & (!type == ("Gilt")))
```

### Model setup
```{r}
model_ermb_ant <- glmer(copies.g ~ type*antibiotic + (1|farm:biorep), data = ermb_ant, family = Gamma(link = "log"))
model_ermb_ant

model_performance(model_ermb_ant)
```

### ermB EMMEANS for antibiotic effects
```{r}

model_ermb_ant_emm <- emmeans(model_ermb_ant, c("antibiotic", "type"))
model_ermb_ant_emm

plot(model_ermb_ant_emm)

model_ermb_ant_emm1 <- emmeans(model_ermb_ant, c("antibiotic"))
model_ermb_ant_emm1

plot(model_ermb_ant_emm1)

model_ermb_ant_emm2 <- emmeans(model_ermb_ant, c("type"))
model_ermb_ant_emm2

joint_tests(model_ermb_ant)

emmip(model_ermb_ant, type ~ antibiotic)
emmip(model_ermb_ant, antibiotic ~ type)

emmeans(model_ermb_ant, pairwise ~ antibiotic | type)
emmeans(model_ermb_ant, pairwise ~ type | antibiotic)
```

## tetM for antibiotic effects
### Data curation
Keeps the tetM gene and removes the integrator 1 and Gilt from the dataset.
```{r}
tetm <- subset(qpcr_r, gene == "tetM")
tetm_ant <- tetm %>%
  filter(integrator == ("2") & !type == ("Gilt"))
```

### Model setup
```{r}
model_tetm_ant <- glmer(copies.g ~ type*antibiotic + (1|farm:biorep), data = tetm_ant, family = Gamma(link = "log"))
model_tetm_ant

model_performance(model_tetm_ant)
```

### tetm EMMEANS for antibiotic effects
```{r}

model_tetm_ant_emm <- emmeans(model_tetm_ant, c("antibiotic", "type"))
model_tetm_ant_emm

plot(model_tetm_ant_emm)

model_tetm_ant_emm1 <- emmeans(model_tetm_ant, c("antibiotic"))
model_tetm_ant_emm1

plot(model_tetm_ant_emm1)

model_tetm_ant_emm2 <- emmeans(model_tetm_ant, c("type"))
model_tetm_ant_emm2

joint_tests(model_tetm_ant)

emmip(model_tetm_ant, type ~ antibiotic)
emmip(model_tetm_ant, antibiotic ~ type)

emmeans(model_tetm_ant, pairwise ~ antibiotic | type)
emmeans(model_tetm_ant, pairwise ~ type | antibiotic)
```

# Soil Versus Manure Comparison with GLMER

## ermB
### Data curation
Keeps the soil and manure, combines it into one datasheet and then filters out 
```{r}
ermb_soil <- subset(ermb, matrix == "soil")
ermb_manure <- subset(ermb, matrix == "manure")
ermb_soil_water <- rbind(ermb_soil, ermb_manure)
```

### Model setup
```{r}
model_ermb_matrix <- glmer(copies.g ~ matrix + (1|farm:biorep), data = ermb_soil_water, family = Gamma(link = "log"))
model_ermb_matrix

model_performance(model_ermb_matrix)
```

### ermb EMMEANS for soil vs manure
```{r}
model_ermb_matrix.emm <- emmeans(model_ermb_matrix, c("matrix"))
model_ermb_matrix.emm

joint_tests(model_ermb_matrix)

plot(model_ermb_matrix.emm)
```

## tetM for soil vs manure
### Data curation
```{r}
tetm_soil <- subset(tetm, matrix == "soil")
tetm_manure <- subset(tetm, matrix == "manure")
tetm_soil_water <- rbind(tetm_soil, tetm_manure)
```

### Model setup
```{r}
model_tetm_matrix <- glmer(copies.g ~ matrix + (1|farm:biorep), data = tetm_soil_water, family = Gamma(link = "log"))
model_tetm_matrix

model_performance(model_tetm_matrix)
```

### tetm EMMEANS for soil vs manure
```{r}
model_tetm_matrix.emm <- emmeans(model_tetm_matrix, c("matrix"))
model_tetm_matrix.emm

joint_tests(model_tetm_matrix)

plot(model_tetm_matrix.emm)
```
